Efficient Unconstraining Parameter Transforms for Hamiltonian Monte Carlo

### 2022-07-17 tex/paper.tex
Update paper.tex

\begin{document}

@@

Sampling algorithms play an important role in Statistics, in this case
we refer to drawing samples from a probability density or distribution
(not survey sampling). Monte Carlo Markov chain methods are commonly
used for high dimensional sampling. In this paper, we look at
Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
samples for the Metropolis algorithm. Bayesian Inference often uses
parameters with constraints, but HMC optimizes on the unconstrained
space. It is difficult to use Monte Carlo estimation on a constrained
domain ~\cite{neal2008optimal}

@@

\subsection{Additive log ratio transform}

@@

The correction term $p_Y(y) = p_X(f^{-1}(y))\,
\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
\subsubsection{Softmax Transform}

@@

\subsection{Softmax Transform}

The softmax function can be understood from Multinomial Logistic
Regression employed for predicting probabilities of a Categorically
distributed variable. Geometrically it maps $R^K$ to the boundary of a
unit K-simplex(it is simply the convex hull of $k+1$ affinely
independent points in $R^K$. Essentially it transforms a vector of
size K to another vector of size K where the outputs sum to 1. It is
worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
when a vector of size $K$ is transformed the $K_th$ vector is simply
$1$- sumof k-1 vectors

>>>>

% transforms -> change of variables
% two kinds of transforms
% pragprg
\begin{document}

@@

%Sampling algorithms play an important role in Statistics, in this case
%we refer to drawing samples from a probability density or distribution
%(not survey sampling). Monte Carlo Markov chain methods are commonly
%used for high dimensional sampling. In this paper, we look at
%Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
%samples for the Metropolis algorithm. Bayesian Inference often uses
%parameters with constraints, but HMC optimizes on the unconstrained
%space. It is difficult to use Monte Carlo estimation on a constrained
%domain ~\cite{neal2008optimal}

@@

@@

The correction term $p_Y(y) = p_X(f^{-1}(y))\,
\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
\subsection{Additive log ratio transform}

\subsubsection{Softmax Transform}

@@

%\subsection{Softmax Transform}
%
%The softmax function can be understood from Multinomial Logistic
%Regression employed for predicting probabilities of a Categorically
%distributed variable. Geometrically it maps $R^K$ to the boundary of a
%unit K-simplex(it is simply the convex hull of $k+1$ affinely
%independent points in $R^K$. Essentially it transforms a vector of
%size K to another vector of size K where the outputs sum to 1. It is
%worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
%when a vector of size $K$ is transformed the $K_th$ vector is simply
%$1$- sumof k-1 vectors
